---
title: "Human Activity Recognition -- Predict Classe"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

<p align="center">[Overview](#overview) | [Analysis](#analysis) | [Modeling](#modeling) |
[Testing](#testing) | [Conclusions](#summary) | [References](#refs)</p>

# <a name="overview" />Overview

In this project, we'll develop a model to predict a discrete variable "classe" (ranking) based on other variables present in the [Weight Lifting Exercises dataset](http://groupware.les.inf.puc-rio.br/har), which ranks an activity performed by the wearer (from A -- "best" to E -- "worst"). The data was collected from from accelerometers on the belt, forearm, arm, and dumbbell of 6 participants.

# <a name="analysis" />Analysis 

## Setup

```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(randomForest)
library(caret)

set.seed(1234)

trainData <- read.csv(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
    na.strings = c("NA","#DIV/0!", "")
)
testData <- read.csv(
    "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
    na.strings = c("NA","#DIV/0!", "")
)
```

## Subsets

```{r}
inTrain <- createDataPartition(trainData$classe, p = 0.7, list = FALSE)
training <- trainData[inTrain,]
testing <- trainData[-inTrain,]
dim(training)
dim(testing)
```


## Clean the Data

```{r}
clear <- grep("name|timestamp|window|X", colnames(training), value = FALSE)
training <- training[, -clear]
testing <- testing[, -clear]
NAs <- apply(
    training, 2, function(x) {sum(is.na(x))}
) / nrow(training)
training <- training[!(NAs > 0.95)] # Remove cols with NA > 95%
testing <- testing[!(NAs > 0.95)]

ncol(training)
ncol(testing)
```

## EDA

```{r warning=FALSE}
ggplot(training, aes(classe)) + 
    geom_histogram(stat = "count", fill = "cadetblue4") + 
    theme_bw() +
    labs(title = "Level of classe in the training set")
```

# <a name="modeling" />Modeling

## PCA

```{r}
preProcess(training[,1:ncol(training)], method = "pca", thresh = .95)
```


```{r message=FALSE, warning=FALSE}
preProc <- preProcess(training[,1:ncol(training)], method = "pca", pcaComp = 27)
trainingPC <- predict(preProc, training[, 1:ncol(training)])
head(trainingPC, n = 2)
```

## Random Forest

```{r message=FALSE, warning=FALSE}
modFit1 <- randomForest(training$classe ~ ., data = trainingPC)
array(importance(modFit1))
```

# <a name="testing" />Cross-validation (Testing)

```{r message=FALSE, warning=FALSE}
testingPC <- predict(preProc, testing[, 1:length(testing)])
confusionMatrix(testing$classe, predict(modFit1, testingPC))$overall
```

## Validation

```{r message=FALSE, warning=FALSE}
validation <- testData[, -clear]
validation <- validation[!(NAs > 0.95)]
validationPC <- predict(preProc, validation[, 1:length(validation)])
validation$classe <- predict(modFit1, validationPC)
```

# <a name="summary" />Conclusions

The overall conclusions for the observations are:
- Out of 160 variables, 53 were useful for modeling
- 5885 out of 13737 observations were used
- The model has a 97.5% overall accuracy
- Sensitivity between 94.5% - 99.2%
- Specificity between 98.9% - 99.7%

### Detailed Statistics

```{r}
confusionMatrix(testing$classe, predict(modFit1, testingPC))$byClass
```

### Final Outcome

```{r}
validation$classe
```


```{r message=FALSE, warning=FALSE}
ggplot(validation, aes(classe)) + geom_histogram(stat = "count", fill = "steelblue2") + theme_bw() +
    labs(title = "Level of classe in the validation set")
```

## <a name="refs" />External References

<i>Ugulino, W.; Cardador, D.; Vega, K.; Velloso, E.; Milidiu, R.; Fuks, H. Wearable Computing: Accelerometers' Data Classification of Body Postures and Movements. Proceedings of 21st Brazilian Symposium on Artificial Intelligence. Advances in Artificial Intelligence - SBIA 2012. In: Lecture Notes in Computer Science. , pp. 52-61. Curitiba, PR: Springer Berlin / Heidelberg, 2012. ISBN 978-3-642-34458-9. DOI: 10.1007/978-3-642-34459-6_6. 
Cited by 2 (Google Scholar)</i>